---
title: "Rapport ARDL"
author: "Koffi Doh ADZAVON"
date: "2023-04-11"
output: html_document
---

The comments and the results of the ARDL model

1-
```{r}

##### read the dataset 
library(tidyverse)
library(zoo)
library(forecast)
library(lubridate)
library(dynlm)
library(nardl)
library(dLagM)
library(dynamic)
library(tseries)
library(TSstudio)
library(dynamac)
library(vars)
library(carData)
library(car)
library(dplyr)
library(ARDL)
library(lmtest)
library(minpack.lm)
library(rgl)
library(qpcR)
library(urca)
setwd("/home/adzavon/Documents/MSC_Data_Analysis/merge for forecasting/")
data <- read_csv("causality_and_var.csv")
data[data == 0] <- 0.001
colnames(data)

```

selection of the variables

```{r}
NV.action <- data$Nvsctions
Vactions <- data$Vactions
Demonstrations <- data$Demonstrations
Floods <- data$Floods
Storms <- data$Storms
Idp <- data$Idp
```

change them into time series data

```{r}
NV.action <- ts(NV.action, start = c(2018,7),
                end = c(2022,12), frequency=12)
Vactions <- ts(Vactions, start = c(2018,7),
                end = c(2022,12), frequency=12)
Demonstrations <- ts(Demonstrations, start = c(2018,7),
                end = c(2022,12), frequency=12)
Floods <- ts(Floods, start = c(2018,7),
                end = c(2022,12), frequency=12)
Storms <- ts(Storms, start = c(2018,7),
                end = c(2022,12), frequency=12)
Idp <- ts(Idp, start = c(2018,7),
                end = c(2022,12), frequency=12)


### metge them to a single time series data

data.ts <- data.frame(Vactions,NV.action,Demonstrations,
                      Floods,Storms,Idp)

```

plot the data

```{r}
plot.ts(data.ts)

```
autocorrelation and partial auto correlation

```{r}

acf(Vactions)
pacf(Vactions)

acf(NV.action)
pacf(NV.action)

acf(Demonstrations)
pacf(Demonstrations)

acf(Floods)
pacf(Floods)

acf(Storms)
pacf(Storms)

acf(Idp)
pacf(Idp)

```
ADF test

```{r}

adf.test(data.ts$Vactions) # not stationary
adf.test(na.omit(data.ts$NV.action))# not stationary
adf.test(data.ts$Demonstrations) # # not stationary
adf.test(data.ts$Floods) # stationary
adf.test(data.ts$Storms) # not stationary
adf.test(data.ts$Idp) # not stationary

```
philip pearon

```{r}

### philip pearon
pp.test(data.ts$Vactions) # stationary
pp.test(data.ts$NV.action) # stationary
pp.test(data.ts$Demonstrations) # stationary
pp.test(data.ts$Floods)  # stationary
pp.test(data.ts$Storms)  # stationary
pp.test(data.ts$Idp) # not stationary

```

differencing the data

```{r}

#  differencing the variables

Vactions.diff <-diff(data.ts$Vactions) # stationary
NV.action.diff <- diff(data.ts$NV.action) # stationary
Demonstrations.diff <- diff(data.ts$Demonstrations) # stationary
Floods.diff <- diff(data.ts$Floods)  # stationary
Storms.diff <- diff(data.ts$Storms)  # stationary
Idp.diff <- diff(data.ts$Idp) # not stationary
```


redo the test of ADF and for Philip pearon


```{r}

# check the stationaruty of the first difference
adf.test(Vactions.diff)
adf.test(NV.action.diff)
adf.test(Demonstrations.diff)
adf.test(Floods.diff)
adf.test(Storms.diff)
adf.test(Idp.diff)

pp.test(Vactions.diff)
pp.test(NV.action.diff)
pp.test(Demonstrations.diff)
pp.test(Floods.diff)
pp.test(Storms.diff)
pp.test(Idp.diff)


```
model to select the max lag

```{r}

max_order_ardl <- auto_ardl(Idp ~ Vactions
                            + NV.action
                            + Demonstrations
                            + Floods
                            + Storms,
                            data = data.ts, selection = "AIC",
                            max_order = 4,
                            selection_minmax = "min"
                           )


max_order_ardl$top_orders
max_order_ardl$best_order
max_order_ardl$best_model
summary(max_order_ardl$best_model)

```

Residual tests

```{r}

res = residuals(max_order_ardl$best_model)
res
mean(res)
jarque.bera.test(res)
shapiro.test(res)
```
my ARDL model for estimation
```{r}
ardl.model <- ardl(Idp ~ Vactions
                            + NV.action
                            + Demonstrations
                            + Floods
                            + Storms,
                            data = data.ts, selection = "AIC",
                            order = c(3,4,3,0,3,3),
                            selection_minmax = "min"
) 

summary(ardl.model)

```

cointegration test

```{r}
bounds_f_test(ardl.model,case = 3, alpha = 0.05)
bounds_t_test(ardl.model, case = 3, alpha = 0.05)
m = ca.jo(data.frame(data.ts), ecdet = "const",type = "trace")
summary(m)
```

unrestricted error correction model

```{r}
# # long run and short run after checking the cointegration
# # uecm (unrestricted error correction model)
ce2_ardl <- coint_eq(ardl.model, case = 3)
uecm <- uecm(ardl.model)
summary(uecm)
```
restricted error corection model

```{r}
# # the shorst run olny (recm restricted error correction model)
ce2_uecm <- coint_eq(uecm, case = 3)
recm <- recm(uecm,case = 3)
summary(recm)
ce2_recm <- coint_eq(recm)
identical(ce2_ardl,ce2_uecm,ce2_recm)
```
long run estimation

```{r}
# # long run estimation

lon_run1 <- multipliers(ardl.model)
lon_run1

long_run2 <- multipliers(uecm)
long_run2
out <- cbind.zoo(LRM = data$Idp, ce2_recm)
plot(out)
```

consum to test the stabitlity of the model
```{r}
par(mar=c(5, 4, 4, 4))
bound.test <- ardlBound(Idp ~ Vactions
                        + NV.action
                        + Demonstrations
                        + Floods
                        + Storms ,
                        data = data.ts, case = 3,
                         max.p = 4, max.q = 2)


```
stability test

```{r}

par(mar = c(5, 5, 2, 2))  # adjust plot margins

bound.test <- ardlBound(Idp ~ Vactions
                        + NV.action
                        + Demonstrations
                        + Floods
                        + Storms ,
                        data = data.ts, case = 3,
                        stability = TRUE, ECM = FALSE,
                        max.p = 4, max.q = 2,
)

```

